---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

```{r include=FALSE}
knitr::opts_chunk$set(warning=FALSE, comment=NA)
```

## Introduction

`darwinator` is an R package that can download sampling event datasets published at GBIF.org from IPT installations.

## Installing from github

If you want to install the latest version of the `darwinator` package from github, you can do it like so:

```{r message=FALSE, warning=FALSE, results="hide", eval=FALSE}
# First make sure you have the devtools package
# which simplifies installations from github
# Note: Windows users have to first install Rtools to use devtools

install.packages("devtools") 
library(devtools)
install_github("GBIF-Europe/darwinator")

```

## Quick start

To see some quick usage examples to get you started, open the Vignette.

Load the package in your R environment:


```{r message=FALSE, warning=FALSE, results="hide", eval=FALSE}

library(darwinator)
sed <- sampling_event_data("78360224-5493-45fd-a9a0-c336557f09c3")

df <- sed$data
dwca <- sed$dwca
citation <- sed$meta$citation

```

Another example - get GBIF dataset identifiers for Norwegian sampling event based datasets using `rgbif`. Then download sampling event data from Norway making use of `purrr::map` and `purrr::possibly`, finally saving to disk.

```{r message=FALSE, warning=FALSE, results="hide", eval=FALSE}
library(tidyverse)
library(purrr)
library(rgbif)
library(darwinator)

search <-
  dataset_search(
    query = "*",
    type = "SAMPLING_EVENT",
    publishingCountry = "NO"
  )

keys <- search$data$datasetKey

# use dplyr::possibly to wrap the sampling_event_data function
# so that it returns NULL if it fails on a particular dataset

psed <- possibly(function(key) sampling_event_data(key), NULL)

norway_samplingeventdata <- map(keys, psed)

saveRDS(norway_samplingeventdata, "no.Rds")

```

## Meta

* Please [report any issues or bugs](https://github.com/GBIF-Europe/darwinator/issues).
* License: AGPL
