---
title: "Introduction to Darwinator"
author: "Markus Skyttner"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Darwinator}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Getting Sample Event Based datasets from GBIF

First search for sampling event datasets from Norway using `rgbif`:

```{r, message=FALSE}

library(rgbif)
library(tidyverse)

search <-
  rgbif::dataset_search(
    type = "SAMPLING_EVENT",
    publishingCountry = "NO"
  )
```

Lets quickly list the first 5 search hits:

```{r}
knitr::kable(
  search$data %>% 
    head(5) %>% 
    select(datasetTitle, publishingOrganization, datasetKey)
  )

```

Now we retrieve specific keys/identifiers for a specific relevant dataset, lets say we want datasets with a title that contains the string "Lepidurus":

```{r}

key <-
  search$data %>%
  filter(grepl("Lepidurus", datasetTitle)) %>%
  .$datasetKey

key

```

This is the dataset we're interested in. Now we retrieve the sampling event data for that specific dataset identifier:

```{r}
library(darwinator)

sed <- sampling_event_data(key)

# we now have a suggested citation to use 
citation <- sed$meta$citation

citation
```

We can display the individual sampling event tables and the combined data from those tables too:

```{r}
library(DT)

# the sample event data tables
event_core <- sed$dwca$event_core
occ_ext <- sed$dwca$occurrence_ext
mof <- sed$dwca$measurementorfact

datatable(event_core %>% head(5))
datatable(occ_ext %>% head(5))
datatable(mof %>% head(5))

```

This is the combined data:

```{r}
# we have the data available for use and the metadata
metadata <- sed$meta
data <- sed$data

# display first 5 rows of data
datatable(
  data %>% 
  head(5)
)

```

